---
title: "Communities"
subtitle: "Movement"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Communities}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
library(ramp.micro) 
```

The matrices describing movement are complex object that we can analyze using methods from network analysis. An R package called *igraph* was developed for this, and there are many other packages that can do this. To take advantage of this functionality, we must first re-encode the graphs in a form. Most of the methods are developed to analyze simple graphs:

## Graphs

+ **`ramp.micro`** includes a function called `make_all_graphs` that creates a graph object for each one of the movement matrices. 

    - **`ramp.micro`** :: `make_graph_obj(M)` is the function that creates and analyzes an adjacency matrix $M$ as a graph:

    - **`igraph`** :: `graph_from_adjacency_matrix(M)-> G` returns a graph $G$ that can be analyzed
    
    - **`igraph`** :: `clusters_walktrap(G)` returns an analysis of the community structure using the walktrap algorithm 

    - **`igraph`** :: `clusters_greedy(M %*% t(M))-> C` returns an analysis of the community structure of the symmetric weighted graph  

These graph objects are all attached by name to `model$graphs.` Note that these analysis are all run by calling **`ramp.micro`** :: `basic_analysis`

```{r}
set.seed(24328)
bb = unif_xy(256, -17, 17) 
qq = unif_xy(289, -17, 17) 
ker_b = make_kF_exp(k=2, s=1, gamma=1.5)
ker_q = make_kF_exp(k=2, s=2, gamma=2)
dispO = list(kFb=ker_b, kFq =ker_q)
bq_mod1 = setup_model(b=bb, q=qq, dispersal_opts = dispO) 
bq_mod1 <- basic_analysis(bq_mod1)
```



```{r, fig.height=4, fig.width=12}
par(mar = c(1,1,4,4), mfrow=c(1,3))
plot_subgraph(bq_mod1, bq_mod1$graphs$GG_graph, stretch=0.1, min_edge_frac=0.1, cut=15, mtl = "Egg Dispersal")
with(bq_mod1, frame_bq(b,q))
add_hulls(bq_mod1, bq_mod1$graphs$GG_graph, stretch=-0.01, min_edge_frac=0.1, cut=15)
add_hulls(bq_mod1, bq_mod1$graphs$VC_graph, stretch=-0.01, min_edge_frac=0.1, cut=15,)
plot_subgraph(bq_mod1, bq_mod1$graphs$VC_graph, stretch=0.1, min_edge_frac=0.1, cut=15, mtl = "Potential Transmission")
```

