---
title: "Models"
subtitle: "Setting Up Models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
library(ramp.micro)
```

## Setup Overview 

**`ramp.micro`** has made it easy to set up models for simulation and analysis. To set up a model: 

1. Define Point Sets (Also, see [Point Sets](point_sets.html))

2. Define Kernel Shapes (Also, see [Kernels](kernels.html)) 

3. Define Parameters. Since **`ramp.micro`** is modular, this must be done for each one of the components: 

    a. Define a module for the Adult component 
    
    b. Define a module for the Aquatic component 

4. Set Initial Conditions. This must be done for each one of the components. The same algorithms can be called to *reset* the initial conditions: 

    a. Set initial conditions for the variables for the Adult component 
    
    b. Set initial conditions for the variables for the Aquatic component 

5. Call Setup     

6. Saving Models
   
## 1. Define Point Sets 

First, we define three point sets: 

```{r}
set.seed(24328)
bb = unif_xy(256, -17, 17) 
qq = unif_xy(289, -17, 17) 
ss = unif_xy(225, -17, 17) 
```

## 2. Define Kernel Shapes 

Next, we define the shapes to weight points by distance:

```{r}
ker_b = make_kF_exp(k=2, s=1, gamma=1.5)
ker_q = make_kF_exp(k=2, s=2, gamma=2)
ker_s = make_kF_exp(k=3, s=2, gamma=2)
```

We pass these as named lists: 

```{r}
bq_dispersal = list(kFb = ker_b, kFq = ker_q)
bqs_dispersal = list(kFb = ker_b, kFq = ker_q, kFs=ker_s)
```

## 3. Define Parameters

The models define default parameter values that can be overwritten by passing alternative values by name. The convention of setting up models by passing named lists makes it makes it easy to get started, and it provides a template that illustrates how to *modify* the parameters, but using the default values can become a trap for lazy analysts. 

### a. Adult Parameters

To see the options for each model, look at the documentation. For the **`BQ`** module:

```{r, eval=F}
?setup_bionomics_BQ
```

For the **`BQS`** module:

```{r, eval=F}
?setup_bionomics_BQS
```

For example, if we wanted to assign random variates drawn from a `beta` distribution to describes survival at sites with a mean of 96%, we would pass the values by name in a list: 

```{r}
adult_opts1 = list(pB = rbeta(256, 96, 4)) 
```

This is used in the examples below. 

Later, when we analyze *potential* transmission, we will need to define the extrinsic incubation period, `eip.`

```{r}
adult_opts2 = list(pB = rbeta(256, 96, 4), eip=12) 
```

### b. Aquatic Parameters

Similarly, to set up parameters for the aquatic habitats, we can set any of the parameters from the command line. For the `basicL` module: 

```{r, eval=F}
?setup_aquatic_model_basicL
```

## 4. Initial Conditions 

To simulate any dynamic model, we must set the initial conditions. Once again, this is done separately for the two components.  

### a. Adult Model 

The current state of the system is stored in `model$Mvars.` 

```{r, eval=F}
?init_adult_model_BQ
```

or 

```{r, eval=F}
?init_adult_model_BQS
```


### b. Aquatic Model 

```{r, eval=F}
?init_aquatic_model_basicL
```

## 5. Setup 

The *setup* functions are designed to create a model object that is ready for simulation and analysis. Documentation for the function `setup_model` explains what it is looking for.  

```{r, eval=F}
?setup_model
```

### Setup BQ 

To set up `BQ,`  we set `Mname = "BQ".` We must pass the point sets `bb` and `qq` and functions that compute kernel weights by distance: `kFb` and `kFq` and `kFs.` 

```{r}
bq_mod1 = setup_model(b=bb, q=qq, kFb=ker_b, kFq=ker_q, bionomic_opts = adult_opts1)
```

### Setup BQS

To set up `BQS,`  we set `Mname = "BQS".` We must pass the point sets `bb` and `qq` and `ss` and functions that compute kernel weights by distance: `kFb` and `kFq` and `kFs.` 

```{r}
bqs_mod1 = setup_model(b=bb, q=qq, s=ss, kFb = ker_b, kFq = ker_q, kFs = ker_s, bionomic_opts = adult_opts2)
```

## 6. Saving & Loading Models

```{r}
save(bq_mod1, file = "bq_mod1.rda")
save(bqs_mod1, file = "bqs_mod1.rda")
```

```{r}
rm("bq_mod1")
exists("bq_mod1")
```



```{r}
load("bq_mod1.rda")
exists("bq_mod1")
```
