---
title: "Building a Model"
subtitle: "A Step-by-Step Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Building a Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
library(ramp.micro)
```

## Setup Overview 

**`ramp.micro`** has made it easy to set up models for simulation and analysis. To set up a model: 

1. Choose the **Modules** (*eg,* `BQ` vs. `BQS` for adult mosquitoes) 

2. Create point sets 

3. Define dispersal kernels 

4. Set parameter values for the **Modules** 

5. Set initial conditions. 

6. Call `setup_model` 

7. Simulate / Analyze

8. Saving Models
   
## 1. Choose the Modules

The first step is to determine which module / model family you want to use for the adult and aquatic components.

**Adult Ecology**

+ `BQ` requires two point sets, $\left\{b\right\}$ and $\left\{q\right\}$ and four dispersal kernels

+ `BQS` requires two point sets, $\left\{b\right\}$ and $\left\{q\right\}$ and $\left\{s\right\}$ and nine dispersal kernels

**Aquatic Ecology**

At the present time, `basicL` is the only module for aquatic ecology, but other modules are planned. 



## 2. Define Point Sets 

**`ramp.micro`** accepts any set of $x,y$ coordinates a user has created.  

Here, we use a **`ramp.micro`** function `unif_xy` to set up 3 sets of points drawn from a random uniform distribution: 

```{r}
set.seed(24328)
bb = unif_xy(256, -17, 17) 
qq = unif_xy(289, -17, 17) 
ss = unif_xy(225, -17, 17) 
```

By convention, each point set is a set of $x,y$ coordinates, and the object is pair of named vectors. To see, we take a peak at $\left\{b\right\}$: 

```{r}
head(bb,3)
```

*See the vignette on [Point Sets](point_sets.html)* 

## 3. Define Kernel Shapes 

Next, we define the shapes to weight points by distance:

```{r}
ker_b = make_kF_exp(k=2, s=1, gamma=1.5)
ker_q = make_kF_exp(k=2, s=2, gamma=2)
ker_s = make_kF_exp(k=3, s=2, gamma=2)
```

Another option is a parameter that determines the fraction of mosquitoes that would stay at a point, called `stayB` and `stayQ` and `stayS` that are set to $0$ if unspecified. If specified, they should be configured by passing either a scalar or a list of values the same length as the number of points.  

We pass these as named lists: 

```{r}
bq_dispersal = list(kFb = ker_b, kFq = ker_q, stayB=0.5, stayQ=0.5)
bqs_dispersal = list(kFb = ker_b, kFq = ker_q, kFs=ker_s)
```

*Also, see the vignette on [Kernels](kernels.html)* 

## 4. Define Parameters

The models define default parameter values that can be overwritten by passing alternative values by name. The convention of setting up models by passing named lists makes it makes it easy to get started, and it provides a template that illustrates how to *modify* the parameters, but using the default values can become a trap for lazy analysts. 

### a. Adult Bionomics 

To see the options for each model, look at the documentation. For the **`BQ`** module:

```{r, eval=F}
?setup_bionomics_BQ
```

For the **`BQS`** module:

```{r, eval=F}
?setup_bionomics_BQS
```

For example, if we wanted to assign random variates drawn from a `beta` distribution to describes survival at sites with a mean of 96%, we would pass the values by name in a list: 

```{r}
adult_opts1 = list(pB = rbeta(256, 96, 4)) 
```

This is used in the examples below. 

Later, when we analyze *potential* transmission, we will need to define the extrinsic incubation period, `eip.`

```{r}
adult_opts2 = list(pB = rbeta(256, 96, 4), eip=12) 
```

### b. Aquatic Parameters

Similarly, to set up parameters for the aquatic habitats, we can set any of the parameters from the command line. For the `basicL` module: 

```{r, eval=F}
?setup_aquatic_model_basicL
```

## 5. Initial Conditions 

To simulate any dynamic model, we must set the initial conditions. Once again, this is done separately for the two components.  

### a. Adult Variables 

The current state of the system is stored in `model$Mvars.` 

```{r, eval=F}
?init_adult_model_BQ
```

or 

```{r, eval=F}
?init_adult_model_BQS
```


### b. Aquatic Variables 

```{r, eval=F}
?init_aquatic_model_basicL
```

## 6. Setup 

The *setup* functions are designed to create a model object that is ready for simulation and analysis. Documentation for the function `setup_model` explains what it is looking for.  

```{r, eval=F}
?setup_model
```

### Setup BQ 

To set up `BQ,`  we set `Mname = "BQ".` We must pass the point sets `bb` and `qq` and functions that compute kernel weights by distance: `kFb` and `kFq` and `kFs.` 

```{r}
bq_mod1 = setup_model(b=bb, q=qq, 
                      kFb=ker_b, kFq=ker_q, 
                      bionomic_opts = adult_opts1)
```

### Setup BQS

To set up `BQS,`  we set `Mname = "BQS".` We must pass the point sets `bb` and `qq` and `ss` and functions that compute kernel weights by distance: `kFb` and `kFq` and `kFs.` 

```{r}
bqs_mod1 = setup_model(b=bb, q=qq, s=ss, 
                       kFb = ker_b, kFq = ker_q, kFs = ker_s, 
                       Mname = "BQS", 
                       bionomic_opts = adult_opts2)
```

## 7. Simulate / Analyze 

```{r}
bq_mod1 <- basic_analysis(bq_mod1)
```

```{r}
bqs_mod1 <- basic_analysis(bqs_mod1)
```


## 8. Saving & Loading Models

```{r}
save(bq_mod1, file = "bq_mod1.rda")
save(bqs_mod1, file = "bqs_mod1.rda")
```

```{r}
rm("bq_mod1")
exists("bq_mod1")
```



```{r}
load("bq_mod1.rda")
exists("bq_mod1")
```
