---
title: "Point Sets"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Point Sets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
library(ramp.micro)
```

## Introduction 

In `ramp.micro`, resource point sets are stored in two columns, and there are utilities for creating point sets for toy models. This is a quick guide. 

## Uniform

`unif_xy(n, mn, mx)` generates random uniform point sets on a box: 

+ `mn` $< x <$ `mx` 

+ `mn` $< y <$ `mx` 

The points are returned as a pair of named vectors -- the first column is named `x` and the next one `y` 

```{r}
p1 <- unif_xy(150, -10, 10)
head(p1, 4)
```
Note that `p1` is $150 \times 2$ 

```{r}
dim(p1)
```

```{r, fig.height=5, fig.width=5}
par(mar = c(2,2,1,1), bty = "o")
plot(p1)
```

## Lattice 

The function `lattice(n, mn, mx)` generates an $n \times n$ lattice inside a box:

+ `mn` $< x <$ `mx` 

+ `mn` $< y <$ `mx`  

```{r, fig.height=5, fig.width=5}
p2 <- lattice(17, -8, 8)
par(mar = c(2,2,1,1), bty = "o")
plot(p2)
```

```{r}
dim(p2)
```

## Clusters 

Two functions were developed to generate clusters. The first one, `clusters_xy(xy, nc, vr)` generates a poisson number of points (mean = `nc`) around a set of seeds. The distance from the center of each seed is a normal variate with variance `vr.` 

```{r}
seeds <- unif_xy(23, -20, 20)
p3 <- clusters_xy(seeds, 8, 1)
```

```{r, fig.height=5, fig.width=5}
par(mar = c(2,2,1,1), bty = "o")
plot(p3)
```
