---
title: "On Dispersal Matrices" 
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{On Dispersal Matrices}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r suppressMessages=T}
library(viridisLite)
library(knitr)
library(viridis)
library(ramp.micro)
#suppressWarnings(devtools::load_all() )
set.seed(25)
```

## Introduction

In behavioral state micro-simulation, mosquito movement is defined by a *search* for a resource, and all resources are located at points in space, $(x,y).$  Micro-dispersal is defined herein as dispersal among point sets representing the locations of resources.

Micro-dispersal models in this software implementation follow a common notation: 

+ A search originates in one point set, $P_v$, with $|v|$ locations; and the destination is another, $P_w$ with $|w|$ locations. 

+ A dispersal model is instantiated as a matrix $M_{w \leftarrow v},$ where the element $M_{i,j} \in M$ is the fraction of mosquitoes moving from $(x_j, y_j) \in P_v$ to $(x_i, y_i) \in P_w$. 

+ If we had a vector, $m_v$, of mosquito abundances at the origin of a search, and if we want to compute a vector, $m_w$,  describing the number arriving at a destination, then we write $$m_w = M \cdot m_v,$$ where

    + $\mbox{dim}(M) = |w| \times |v|,$ 
    
    + and $|m_v| = |v|,$ 
    
    + so $|m_w| = |w|.$ 


By convention, we constrain model so that, $$\sum_i M_{i,j}=1.$$  

If we want to model failed dispersal, then we will often want to handle the case where the origin and destination are the same point set. 

In the following, we discuss the challenge of visualizing micro-dispersal matrices using functions in `motrap.micro` where either $P_v \neq P_w$  or where $P_v = P_w.$ 

##  Dispersal from $P_v \neq P_w$ 

To illustrate, we generate two point sets:

```{r}
nPv = 15 
nPw = 7
dd = 10
Pv = unif_xy(nPv,-dd, dd) 
Pw = unif_xy(nPw,-dd, dd) 
```

Now, we can define a dispersal matrix from $S$ to $D$:

```{r}
M_wv = make_Psi_xy(Pv, Pw)
```

The generic function `plot_Mxy` was developed to plot dispersal from $P_v$ to $P_w.$ 

```{r, fig.height=5, fig.width=5, echo=F}
par(mar = c(1,1,1,1))
plot_matrix_xy(Pv, Pw, M_wv, r=0.25) -> wts
```

```{r}
M_vw = make_Psi_xy(Pw, Pv)
Mvv = M_vw %*% M_wv
Mww = M_wv %*% M_vw
```


```{r, fig.height=5, fig.width=10, echo=F}
par(mar = c(1,1,1,1), mfcol = c(1,2))
plot_matrix_xx(Pv, Mvv, min_edge_frac=1e-3, lwd=23, r=0.25) -> wts
plot_matrix_xx(Pw, Mww, min_edge_frac=1e-3, lwd=23, r=0.25) -> wts
```
